---
title: "Evaluation and the causal revolution"
linktitle: "1: Evaluation and the causal revolution"
date: "2020-08-24"
start_date: "2020-08-24"
end_date: "2020-08-28"
menu:
  content:
    parent: Course content
    weight: 1
type: docs
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-syllabus-no-bib.csl"
slides: "01-slides"
output:
  blogdown::html_page:
    toc: true
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-tabs.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sHcnk9KcoQVzDywxWRzVWHC"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "BIhz1dz5Vws", "Introduction", "1", TRUE,
  "f-sOSXyeeJE", "Data science and public service", "ds-ps", FALSE,
  "Z2foqSMWnGU", "Evidence, evaluation, and causation (1)", "evidence-evaluation-causation", FALSE,
  "W5ChuEhAzrs", "Evidence, evaluation, and causation (2)", "evidence-evaluation-causation1", FALSE,
  "D4LoEctZxHE", "Class details", "class-details", FALSE
)
```

## Readings

- The [syllabus](/syllabus/), [content](/content/), [examples](/example/), and [assignments](/assigment/) pages for this class
- <i class="fas fa-book"></i> Chapter 1 in *Impact Evaluation in Practice* [@wb-eval]
- <i class="fab fa-youtube"></i> DJ Patil, ["What Makes a Radical and Revolutionary Technology?"](https://www.youtube.com/watch?v=UOoMzaWOQJA) 
    - <small>(DJ Patil is the former Chief Data Scientist of the United States under President Obama. He gave this forum address at Brigham Young University on February 13, 2018.)</small>
- <i class="fas fa-external-link-square-alt"></i> Stephen Goldsmith, ["Next Generation of Public Employees Must Understand Data and Policy"](http://www.govtech.com/opinion/Next-Generation-Public-Employees-Must-Understand-Data-Policy.html)
- <i class="fas fa-external-link-square-alt"></i> Hadley Wickham, ["Data Science: How is it Different To Statistics?"](http://bulletin.imstat.org/2014/09/data-science-how-is-it-different-to-statistics%E2%80%89/)


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

`r blogdown::shortcode("slide-buttons")`

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_tabs(slide_details, "/slides/01-slides.html")
```

:::fyi
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```

## Optional lab

Here are all the materials we used in the lab: 

- [CSV of the cars dataset](/data/cars.csv)
- [RStudio.cloud project](https://rstudio.cloud/spaces/82624/project/1556360)
- [Slides](https://evalf20.classes.andrewheiss.com/projects/01_lab/slides/01_lab.html)
- [Project `.zip` file](https://evalf20.classes.andrewheiss.com/projects/01_lab.zip) (the slides are in here too)


```{cat include=FALSE, eval=FALSE}
Playlist: https://www.youtube.com/playlist?list=PLS6tnpTr39sHcnk9KcoQVzDywxWRzVWHC

https://youtu.be/BIhz1dz5Vws
PMAP 8521 • (1) Evaluation and the causal revolution: (0) Introduction

Introduction to the class session on evaluation and the causal revolution

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/01-content/


https://youtu.be/f-sOSXyeeJE
PMAP 8521 • (1) Evaluation and the causal revolution: (1) Data science and public service

Discussion of the role of data science in the public and nonprofit sectors

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/01-content/

  
https://youtu.be/Z2foqSMWnGU
PMAP 8521 • (1) Evaluation and the causal revolution: (2.1) Evidence, evaluation, and causation

Discussion of the role of evidence-based policy and program evaluation

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/01-content/


https://youtu.be/W5ChuEhAzrs
PMAP 8521 • (1) Evaluation and the causal revolution: (2.2) Evidence, evaluation, and causation

Discussion of why causation is tricky (but also possible!) to measure

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/01-content/


https://youtu.be/D4LoEctZxHE
PMAP 8521 • (1) Evaluation and the causal revolution: (3) Class details

Overview of how the class works

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/01-content/

```
