---
title: "Regression and inference"
linktitle: "2: Regression and inference"
date: "2020-08-31"
start_date: "2020-08-31"
end_date: "2020-09-04"
menu:
  content:
    parent: Course content
    weight: 2
type: docs
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-syllabus-no-bib.csl"
slides: "02-slides"
output:
  blogdown::html_page:
    toc: true
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-tabs.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sERjvBbJGr9mpqvtiLVPzd4"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "S4OmWxc3h3Y", "Introduction", "1", TRUE,
  "DGbLEgL4g6Y", "Drawing lines", "drawing-lines", FALSE,
  "n-rV1TIoIgw", "Lines, Greek, and regression", "lines-greek-regression", FALSE,
  "RYKsrTKWCR4", "Null worlds and statistical significance", "significance", FALSE
)
```

## Readings

- <i class="fas fa-book"></i> Chapter 2 in *Mastering 'Metrics* [@AngristPischke:2015]
- <i class="fas fa-book"></i> "Properties of regression" in *Causal Inference: The Mixtape* [@mixtape]

## Recommended readings

Look through your notes on regression from your last stats class. Also, you can skim through these resources:

- <i class="fas fa-book"></i> [6.1–6.4 in *ModernDive*](https://moderndive.com/6-regression.html) [@IsmayKim:2018]
- <i class="fas fa-book"></i> [7.1–7.4 in *ModernDive*](https://moderndive.com/7-multiple-regression.html) [@IsmayKim:2018]
- <i class="fas fa-book"></i> 7.1–7.3 in *OpenIntro Statistics* [@DiezBarrCetinkaya-Rundel:2017]
- <i class="fas fa-book"></i> 8.1 in *OpenIntro Statistics* [@DiezBarrCetinkaya-Rundel:2017]

We'll review all this regression stuff in the videos, so *don't panic* if this all looks terrifying! Also, take advantage of the videos that accompany the *OpenIntro* chapters. And *also*, the *OpenIntro* chapters are heavier on the math—don't worry if you don't understand everything.


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

`r blogdown::shortcode("slide-buttons")`

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_tabs(slide_details, "/slides/02-slides.html")
```

:::fyi
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```

```{cat include=FALSE, eval=FALSE}
Playlist: https://www.youtube.com/playlist?list=PLS6tnpTr39sERjvBbJGr9mpqvtiLVPzd4

https://youtu.be/S4OmWxc3h3Y
PMAP 8521 • (2) Regression and inference: (0) Introduction

Introduction to the class session on regression and inference

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/02-content/


https://youtu.be/DGbLEgL4g6Y
PMAP 8521 • (2) Regression and inference: (1) Drawing lines

Overview of the basics of regression analysis (particularly ordinary least squares (OLS) regression)

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/02-content/

  
https://youtu.be/n-rV1TIoIgw
PMAP 8521 • (2) Regression and inference: (2) Lines, Greek, and regression

Discussion of how to estimate parameters from a population using regression (with R specifically)

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/02-content/


https://youtu.be/RYKsrTKWCR4
PMAP 8521 • (2) Regression and inference: (3) Null worlds and statistical significance

Discussion of statistical significance and p-values using simulated null worlds

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/02-content/


https://youtu.be/E-Zz5S5NOUo
PMAP 8521 • Example: Regression with R

Demonstration of how to run regression models with R

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/regression/
```
