---
title: "Threats to validity"
linktitle: "6: Threats to validity"
date: "2020-09-28"
start_date: "2020-09-28"
end_date: "2020-10-02"
menu:
  content:
    parent: Course content
    weight: 6
type: docs
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-syllabus-no-bib.csl"
slides: "06-slides"
output:
  blogdown::html_page:
    toc: true
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-tabs.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sFJtz7egfUASfhhz6zuj7GZ"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "46v8l60nQnI", "Introduction", "1", TRUE,
  "N4XYXuw_IA0", "Construct validity", "construct-validity", FALSE,
  "IziQBDzUJ6E", "Statistical conclusion validity", "statistical-conclusion-validity", FALSE,
  "ssz2DTfrGhw", "Internal validity", "internal-validity", FALSE,
  "uK7YN6KTRT4", "External validity", "external-validity", FALSE
)
```


## Readings

- <i class="fas fa-external-link-square-alt"></i> Randall Munroe, ["Significant"](https://xkcd.com/882/)
- <i class="fas fa-external-link-square-alt"></i> Alexander Coppock, ["10 Things to Know About Statistical Power"](https://egap.org/resource/10-things-to-know-about-statistical-power/)
- <i class="fas fa-trophy"></i> Play around with FiveThirtyEight, ["Hack Your Way To Scientific Glory"](https://projects.fivethirtyeight.com/p-hacking/)
- <i class="fas fa-book"></i> Chapter 9 in *Impact Evaluation in Practice* [@wb-eval]


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

`r blogdown::shortcode("slide-buttons")`

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_tabs(slide_details, "/slides/06-slides.html")
```

:::fyi
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```


```{cat include=FALSE, eval=FALSE}
Playlist: https://www.youtube.com/playlist?list=PLS6tnpTr39sFJtz7egfUASfhhz6zuj7GZ

https://youtu.be/46v8l60nQnI
PMAP 8521 • (6) Threats to validity: (0) Introduction

Introduction to the class session on threats to validity

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/06-content/


https://youtu.be/N4XYXuw_IA0
PMAP 8521 • (6) Threats to validity: (1) Construct validity

Discussion of threats to construct validity

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/06-content/

  
https://youtu.be/IziQBDzUJ6E
PMAP 8521 • (6) Threats to validity: (2) Statistical conclusion validity

Discussion of threats to statistical conclusion validity

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/06-content/
  
  
https://youtu.be/ssz2DTfrGhw
PMAP 8521 • (6) Threats to validity: (3) Internal validity

Discussion of threats to internal validity

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/06-content/
  
  
https://youtu.be/uK7YN6KTRT4
PMAP 8521 • (6) Threats to validity: (4) External validity

Discussion of threats to external validity

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/06-content/
```
