---
title: "Regression discontinuity I"
linktitle: "10: RDD I"
date: "2020-10-26"
start_date: "2020-10-26"
end_date: "2020-10-30"
menu:
  content:
    parent: Course content
    weight: 10
type: docs
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-syllabus-no-bib.csl"
slides: "10-slides"
output:
  blogdown::html_page:
    toc: true
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-tabs.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sGYRrLn6OEoI5kry5gvJvdz"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "4y84N23Mx-o", "Introduction", "1", TRUE,
  "ZHBZtKL7zAg", "Arbitrary cutoffs and causal inference", "arbitrary-cutoffs", FALSE,
  "B060u1zGOWo", "Drawing lines and measuring gaps", "lines-gaps", FALSE,
  "0v1aE70FhsQ", "Main RDD concerns", "main-concerns", FALSE
)
```

## Readings

- <i class="fas fa-book"></i> Chapter 6 in *Impact Evaluation in Practice* [@wb-eval]
- <i class="fas fa-book"></i> Chapter 4 in *Mastering 'Metrics* [@AngristPischke:2015]
- <i class="fas fa-book"></i> "Regression discontinuity" in *Causal Inference: The Mixtape* [@mixtape]

### Regression discontinuity

- The [example page on regression discontinuity](/example/rdd/) shows how to use R to analyze and estimate causal effects with regression discontinuity


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

`r blogdown::shortcode("slide-buttons")`

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_tabs(slide_details, "/slides/10-slides.html")
```

:::fyi
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```


```{cat include=FALSE, eval=FALSE}
Playlist: https://www.youtube.com/playlist?list=PLS6tnpTr39sGYRrLn6OEoI5kry5gvJvdz

https://youtu.be/EqSMG277wVI
PMAP 8521 • (10) Regression discontinuity I: (0) Introduction

Introduction to the class session on regression discontinuity designs

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/10-content/


https://youtu.be/vBDNcPHAGfg
PMAP 8521 • (10) Regression discontinuity I: (1) Arbitrary cutoffs and causal inference

Discussion of how arbitrary rules and cutoffs let us generate pseudo-treatment and control groups and find causal effects in observational data

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/10-content/

  
https://youtu.be/ZijPafzc-wY
PMAP 8521 • (10) Regression discontinuity I: (2) Drawing lines and measuring gaps

Discussion of how to draw lines in different ways around a discontinuity and how that influences the size of the gap

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/10-content/
  
  
https://youtu.be/Or-VUzCZIBs
PMAP 8521 • (10) Regression discontinuity I: (3) Main RDD concerns

Discussion of the main assumptions for regression discontinuity design

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/10-content/
  
  
https://www.youtube.com/playlist?list=PLS6tnpTr39sFCr8amjT2Ld_FE34i26a3S

https://youtu.be/8CayGMcJoLY
PMAP 8521 • Example: Regression discontinuity with R: 1: Getting started

Demonstration of how to create a new RStudio project and get ready to analyze regression discontinuity data

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/rdd/


https://youtu.be/ViDChprm2IY
PMAP 8521 • Example: Regression discontinuity with R: 2: Sharp or fuzzy?

Demonstration of how to determine if a cutoff is sharp or fuzzy

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/rdd/


https://youtu.be/hqPVIt-72fY
PMAP 8521 • Example: Regression discontinuity with R: 3: Checking for manipulation

Demonstration of how to check for manipulation in the running variable near the cutpoint with a McCrary density test

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/rdd/


https://youtu.be/yYqO22l_jgc
PMAP 8521 • Example: Regression discontinuity with R: 4: Finding discontinuity in outcome

Demonstration of how to plot a discontinuity in an outcome based on a running variable and a cutpoint

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/rdd/


https://youtu.be/Por8Rd9QuBk
PMAP 8521 • Example: Regression discontinuity with R: 5: Measuring the gap parametrically

Demonstration of how to measure the gap at a discontinuity using parametric methods

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/rdd/
  

https://youtu.be/lfBkAZLX_Z4
PMAP 8521 • Example: Regression discontinuity with R: 6: Measuring the gap nonparametrically

Demonstration of how to measure the gap at a discontinuity using nonparametric methods with rdrobust

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/rdd/
```