---
title: "Instrumental variables I"
linktitle: "11: IV I"
date: "2020-11-02"
start_date: "2020-11-02"
end_date: "2020-11-06"
menu:
  content:
    parent: Course content
    weight: 11
type: docs
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-syllabus-no-bib.csl"
slides: "11-slides"
output:
  blogdown::html_page:
    toc: true
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-tabs.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sEsOSeMK19h6RWUB5kHrn0q"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "2L4h_XRrNAg", "Introduction", "1", TRUE,
  "fLJKAMSXDyw", "Endogeneity and exogeneity", "endo-exo", FALSE,
  "915d3OMaWNA", "Instruments", "instruments", FALSE,
  "rCMQkrVIKOc", "Using instruments", "using-instruments", FALSE
)
```

## Readings

- <i class="fas fa-book"></i> Chapter 5 in *Impact Evaluation in Practice* [@wb-eval]
- <i class="fas fa-book"></i> Chapter 3 in *Mastering 'Metrics* [@AngristPischke:2015]
- <i class="fas fa-book"></i> "Instrumental variables" in *Causal Inference: The Mixtape* [@mixtape]

### Instrumental variables

- The [example page on instrumental variables](/example/iv/) shows how to use R to analyze and estimate causal effects with instrumental variables


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

`r blogdown::shortcode("slide-buttons")`

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_tabs(slide_details, "/slides/11-slides.html")
```

:::fyi
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```


```{r create-save-data, include=FALSE}
library(tidyverse)
library(scales)

set.seed(123456)

nrows <- 1000
father_education <- tibble(
  ability = rnorm(nrows, 35000, 10000),  # Ability
  fathereduc = rnorm(nrows, 15000, 20000),  # Father's education (IV)
  e_y = 0.43 * rnorm(nrows, 50000, 10000)  # Error for outcome
) %>% 
  mutate(educ = 3.7 + 0.52*fathereduc + 0.40*ability,  # Education (policy variable)
         wage = 5 + 0.23*educ + 0.5*ability - e_y) %>%  # Wage (outcome variable)
  mutate(wage = rescale(wage, to = c(7.75, 300)),  # Rescale from minimum wage to director wage (hourly)
         educ = rescale(educ, to = c(10, 23)),  # Rescale as years of school. Min 10 to max 23 (PhD)
         fathereduc = rescale(fathereduc, to = c(10, 23)),  # Rescale father's education
         ability = rescale(ability, to = c(0, 600))) %>%  # Rescale as Hypothetical test scores
  select(wage, educ, ability, fathereduc) %>% 
  mutate(across(everything(), ~round(., 2)))
write_csv(father_education, here::here("static", "data", "father_education.csv"))

write_csv(wooldridge::wage2, here::here("static", "data", "wage2.csv"))

write_csv(wooldridge::card, here::here("static", "data", "card.csv"))
```

```{cat include=FALSE, eval=FALSE}
Playlist: https://www.youtube.com/playlist?list=PLS6tnpTr39sEsOSeMK19h6RWUB5kHrn0q

https://youtu.be/2L4h_XRrNAg
PMAP 8521 • (11) Instrumental variables I: (0) Introduction

Introduction to the class session on instrumental variables

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/11-content/


https://youtu.be/fLJKAMSXDyw
PMAP 8521 • (11) Instrumental variables I: (1) Endogeneity and exogeneity

Discussion of the difference between endogeneity and exogeneity in a DAG

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/11-content/

  
https://youtu.be/915d3OMaWNA
PMAP 8521 • (11) Instrumental variables I: (2) Instruments

Discussion of what an instrument is and how it needs to be relevant, exclusive, and exogenous

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/11-content/
  
  
https://youtu.be/rCMQkrVIKOc
PMAP 8521 • (11) Instrumental variables I: (3) Using instruments

Discussion of how to use instruments in two-stage least squares (2SLS) models to isolate exogeneity in treatment variables

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/11-content/
  
  
https://www.youtube.com/playlist?list=PLS6tnpTr39sF97TfofFZ1O-bOQtB4ygZQ

https://youtu.be/h-rx-dXo74Y
PMAP 8521 • Example: Instrumental variables with R: 1: Getting started

Demonstration of how to create a new RStudio project and get ready to analyze data with instrumental variables

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/iv/


https://youtu.be/wHTiAbGwKe4
PMAP 8521 • Example: Instrumental variables with R: 2: Checking instrument validity

Demonstration of how to check the validity of an instrument using fake data (wages ~ education | father's education)

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/iv/


https://youtu.be/tnerIh0uMxU
PMAP 8521 • Example: Instrumental variables with R: 3: Manual 2SLS

Demonstration of how to run a two-stage least squares (2SLS) regression model by running a first stage model, generated predicted values of the treatment variable, and use those exogenous predictions in a second stage model (using fake data (wages ~ education | father's education))

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/iv/


https://youtu.be/EI-knWTbH-s
PMAP 8521 • Example: Instrumental variables with R: 4: One-step 2SLS

Demonstration of how to run two-stage least squares (2SLS) regression model in one step using the iv_robust function from the **estimatr** R package (using fake data (wages ~ education | father's education))

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/iv/


https://youtu.be/W8Up5OOPtEY
PMAP 8521 • Example: Instrumental variables with R: 5: Checking instrument validity (real data)

Demonstration of how to check instrument validity using real data (wages ~ education | father's education + mother's education)

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/iv/
  

https://youtu.be/6sp8sYrRv5c
PMAP 8521 • Example: Instrumental variables with R: 6: 2SLS (real data)

Demonstration of how to run two-stage least squares (2SLS) regression with multiple instruments, using real data (wages ~ education | father's education + mother's education)

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/iv/


https://youtu.be/uDr4nsWkG2c
PMAP 8521 • Example: Instrumental variables with R: 7: Validity + 2SLS (real data)

Demonstration of how to check instrument validity and run two-stage least squares (2SLS) regression with control variables, using real data (wages ~ education + controls | distance to college + controls)

Download the data and see the polished code at https://evalf20.classes.andrewheiss.com/example/iv/
```
