---
title: "Instrumental variables II + Regression discontinuity II"
linktitle: "12: IV I & II"
date: "2020-11-09"
start_date: "2020-11-09"
end_date: "2020-11-13"
menu:
  content:
    parent: Course content
    weight: 12
type: docs
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-syllabus-no-bib.csl"
slides: "12-slides"
output:
  blogdown::html_page:
    toc: true
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-tabs.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sGrbGfvdp0Gob6qAghIdiLq"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "qDAFXXriFyE", "Introduction", "1", TRUE,
  "_Ue7blCyEXk", "Treatment effects and compliance", "treatment-effects", FALSE,
  "M5cuBBuqZys", "Randomized promotion", "encouragement", FALSE,
  "9xGeCGRvVgE", "Fuzzy regression discontinuity", "fuzzy-rdd", FALSE
)
```

## Readings

- The [example page on complier average causal effects](/example/cace/) shows how to use R to disentangle complier average causal effects
- The [example page on fuzzy regression discontinuity](/example/rdd-fuzzy) shows how to use R to use instrumental variables in fuzzy regression discontinuity, both parametrically and nonparametrically


## Slides

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

`r blogdown::shortcode("slide-buttons")`

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_tabs(slide_details, "/slides/12-slides.html")
```

:::fyi
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

```{r show-youtube-list, echo=FALSE, results="asis"}
youtube_list(slide_details, playlist_id)
```


```{r create-save-data, include=FALSE}
library(tidyverse)
library(scales)

# Fake program data
set.seed(1234)
num_students <- 1000
tutoring <- tibble(
  id = 1:num_students,
  entrance_exam = rbeta(num_students, shape1 = 7, shape2 = 2),
  exit_exam_base = rbeta(num_students, shape1 = 5, shape2 = 3)
) %>% 
  mutate(entrance_exam = entrance_exam * 100) %>% 
  mutate(tutoring_sharp = entrance_exam <= 70) %>%
  mutate(tutoring_fuzzy = case_when(
    entrance_exam >= 50 & entrance_exam <= 70 ~ sample(c(TRUE, FALSE), n(), replace = TRUE, prob = c(0.8, 0.2)),
    entrance_exam > 70 & entrance_exam <= 90 ~ sample(c(TRUE, FALSE), n(), replace = TRUE, prob = c(0.2, 0.8)),
    entrance_exam < 50 ~ TRUE,
    entrance_exam > 90 ~ FALSE
  )) %>% 
  mutate(tutoring_text = factor(tutoring_sharp, levels = c(FALSE, TRUE),
                                labels = c("No tutor", "Tutor")),
         tutoring_fuzzy_text = factor(tutoring_fuzzy, levels = c(FALSE, TRUE),
                                      labels = c("No tutor", "Tutor"))) %>% 
  mutate(exit_exam = exit_exam_base * 40 + 10 * tutoring_sharp + entrance_exam / 2) %>% 
  mutate(exit_exam_fuzzy = exit_exam_base * 40 + 10 * tutoring_fuzzy + entrance_exam / 2) %>% 
  mutate(entrance_centered = entrance_exam - 70) %>% 
  mutate(below_cutoff = entrance_exam <= 70)

tutoring <- tutoring %>% 
  select(id, entrance_exam, tutoring = tutoring_fuzzy, 
         tutoring_text = tutoring_fuzzy_text, exit_exam = exit_exam_fuzzy)

write_csv(tutoring, here::here("static", "data", "tutoring_program_fuzzy.csv"))


set.seed(1234)

N <- 2000
df <- tibble(
  status = sample(c("Always taker", "Never taker", "Complier"), N, 
                  replace = TRUE, prob = c(0.2, 0.4, 0.4)),
  treatment = sample(c("Treatment", "Control"), N, replace = TRUE, prob = c(0.5, 0.5))
) %>% 
  mutate(bed_net_0 = (status == "Always taker") * 1,
         bed_net_1 = (status != "Never taker") * 1) %>% 
  mutate(health_0 = case_when(
    status == "Always taker" ~ rnorm(N, 1, 0.5),
    status == "Never taker"  ~ rnorm(N, 0, 0.6),
    status == "Complier"     ~ rnorm(N, 0.1, 0.4),
  )) %>% 
  mutate(health_1 = case_when(
    status == "Always taker" ~ rnorm(N, 1, 0.5),
    status == "Never taker"  ~ rnorm(N, 0, 0.6),
    status == "Complier"     ~ rnorm(N, 0.9, 0.7),
  )) %>% 
  mutate(bed_net = case_when(
    treatment == "Treatment" ~ bed_net_1,
    treatment == "Control"   ~ bed_net_0
  )) %>% 
  mutate(health = case_when(
    bed_net == 0 ~ health_0,
    bed_net == 1 ~ health_1
  )) %>% 
  mutate(bed_net = factor(bed_net, labels = c("No bed net", "Bed net")))

min_health <- min(df$health_0, df$health_1)
max_health <- max(df$health_0, df$health_1)

df <- df %>% 
  mutate(across(starts_with("health"), 
                ~rescale(., to = c(0, 100), 
                         from = c(min_health, max_health))))

df_real <- df %>% 
  select(treatment, bed_net, health)

write_csv(df, here::here("static", "data", "bed_nets_time_machine.csv"))
write_csv(df_real, here::here("static", "data", "bed_nets_observed.csv"))
```

```{cat include=FALSE, eval=FALSE}
Playlist: https://www.youtube.com/playlist?list=PLS6tnpTr39sGrbGfvdp0Gob6qAghIdiLq

https://youtu.be/qDAFXXriFyE
PMAP 8521 • (12) Instrumental variables II + Regression discontinuity II: (0) Introduction

Introduction to the class session on instrumental variables and fuzzy regression discontinuity

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/12-content/


https://youtu.be/_Ue7blCyEXk
PMAP 8521 • (12) Instrumental variables II + Regression discontinuity II: (1) Treatment effects and compliance

Discussion of compliers, always takers, and never takers + how to find complier average causal effects (CACE) by finding the intent-to-treat effect (ITT) and dividing it by the proportion of compliers

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/12-content/

  
https://youtu.be/M5cuBBuqZys
PMAP 8521 • (12) Instrumental variables II + Regression discontinuity II: (2) Randomized promotion

Discussion of how to use randomized promotion as an instrument to measure the causal effect of a universal program with self-selection

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/12-content/
  
  
https://youtu.be/9xGeCGRvVgE
PMAP 8521 • (12) Instrumental variables II + Regression discontinuity II: (3) Fuzzy regression discontinuity

Discussion of how to use an above/below threshold variable as an instrument in fuzzy regression discontinuity designs

Access the slides and other materials at the course website at https://evalf20.classes.andrewheiss.com/content/12-content/
```
