---
title: "Lab 1 stuff"
output: html_document
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)  # For ggplot, dplyr, and friends

mpg <- read_csv("data/mpg.csv")  # Load data
```

## Figuring out what's in the data

```{r}
mpg
```

## Basic ggplot code

Run this code from the the slide to make a graph.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

## ggplot playground

Add `color`, `size`, `alpha`, or `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Switching geoms

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = hwy))
```

## Histograms

Make a histogram of the `hwy` variable from `mpg`.

```{r}
ggplot(data = mpg)
```

## Density plots

Make a density plot `hwy` colored by `class`.

```{r}
ggplot(data = mpg) 
```

## Complex plots

Predict what this code will do. Then run it.

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))
```

---

## Fancy completed plot from slides

```{r message=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  scale_color_viridis_d() +
  facet_wrap(vars(drv), ncol = 1) +
  labs(x = "Displacement", y = "Highway MPG",
       color = "Drive", 
       title = "Heavier cars get lower mileage",
       subtitle = "Displacement indicates weight(?)",
       caption = "I know nothing about cars") +
  theme_bw() +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold"))
```

---

## Takeaways

You can use this code template to any kind of graph with **ggplot2**.

```{r eval=FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

---

# Manipulating data with dplyr

```{r}
gapminder <- read_csv("data/gapminder.csv")
```

## gapminder

```{r}
gapminder
```


## Your turn 1

- `filter()` selects rows based on logical tests

The data for Canada:

```{r}
filter(gapminder, country == "New Zealand")
```

All data for countries in Oceania:

```{r}
filter(gapminder, country == "New Zealand")
```

Rows where the life expectancy is greater than 82:

```{r}
filter(gapminder, country == "New Zealand")
```


## Your turn 2

Canada before 1970:

```{r}
filter(gapminder, country == "New Zealand", year > 2000)
```

Countries where life expectancy in 2007 is below 50:

```{r}
filter(gapminder, country == "New Zealand", year > 2000)
```

Countries where life expectancy in 2007 is below 50, and are not in Africa:

```{r}
filter(gapminder, country == "New Zealand", year > 2000)
```


## Your turn 3

Write code to:

- Add an `africa` column that is TRUE if the country is on the African continent
- Add a column for logged GDP per capita (hint: use `log()`)
- Add an `africa_asia` column that says “Africa or Asia” if the country is in Africa or Asia, and “Not Africa or Asia” if it’s not

```{r}
mutate(gapminder)
```


## Your turn 4

Use `summarize()` to compute four statistics about the data:

- The first (minimum) year in the dataset
- The last (maximum) year in the dataset
- The number of rows in the dataset (use the cheatsheet)
- The number of distinct countries in the dataset (use the cheatsheet)

```{r}
gapminder 
```


## Your turn 5

Use `filter()` and `summarize()` to calculate the (1) the number of unique countries and (2) the median life expectancy on the African continent in 2007.

```{r}
gapminder 
```


## Your turn 6

Find the minimum, maximum, and median life expectancy for each continent.

```{r}
gapminder 
```

Find the minimum, maximum, and median life expectancy for each continent in 2007 only.

```{r}
gapminder
```
